### 2014-09-26 tex/proposal_letter.tex
Finished latest draft and finalized proposal letter

\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}

\makeatletter
\renewcommand{\closing}[1]{\par\nobreak\vspace{\parskip}%
  \stopbreaks
  \noindent
  \ifx\@empty\fromaddress\else
  \hspace*{\longindentation}\fi
  \parbox{\indentedwidth}{\raggedright
       \ignorespaces #1\\[1\medskipamount]%
       \ifx\@empty\fromsig
           \fromname
       \else \fromsig \fi\strut}%
   \par}
\makeatother

\usepackage{fontspec} 
\defaultfontfeatures{Mapping=tex-text} % converts LaTeX specials (``quotes'' --- dashes etc.) to unicode
\newopentypefeature{Contextuals}{NoAlternate}{-calt} % enable turning off the crazy Q in Garamond Premier Pro
\usepackage{xunicode}
\setmainfont{Garamond Premier Pro}
\setsansfont[Scale=0.9]{Optima Regular}

\usepackage{wallpaper}

\signature{Matthew V. Talluto, \\
	Isabelle Boulangeat, \& \\
	Dominique Gravel \\
	Université du Québec à Rimouski \\ 
	mtalluto@gmail.com}
\address{Matthew V. Talluto, Ph.D. \\
	Départament du Biologie, Chimie, et Géographie \\
	300, Allée des Ursulines \\
	Rimouski, Québec G5L 3A1, Canada \\
	mtalluto@gmail.com}

\begin{document}
\addfontfeatures{Contextuals=NoAlternate}   % turn off the crazy Q in Garamond Premier Pro
\ThisULCornerWallPaper{0.22}{uqar-logo2}

\begin{letter}{Jordi Bascompte \\ 
	Ideas \& Perspectives Editor \\
	\emph{Ecology Letters}
}
\opening{Dear Dr. Bascompte,}

%TC:break opening
We are pleased to submit a proposal for a manuscript titled ``Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework''. 
We would appreciate your consideration of our manuscript for the \emph{Ideas and Perspectives} section of \emph{Ecology Letters}. Please find our proposal text below.

%TC:break body
We present a novel framework for integrating small-scale, more mechanistic models into large-scale correlative species distribution models in order to improve the forecasts of these models.
The processes generating species ranges are complex and operate at multiple scales.
Consequently, a wealth of theory (e.g., metapopulation theory, niche theory, and complexity theory) underlies models of species ranges, resulting in a diversity of approaches.
In practice, however, most models do not include all available information; for example, correlative models predicting presence/absence as a function of climate often ignore smaller-scale processes such as growth, fecundity, and dispersal.
Furthermore, individual models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
Our contribution is the first to propose a conceptual and analytical framework that goes further than \emph{a posteriori} aggregation of predictions from different models. 
We use hierarchical Bayesian methods to integrate knowledge at varying ecological scales in order to provide more robust predictions future species distributions.
Moreover, our method explicitly quantifies uncertainty and propagates it from all information sources.
This is of crucial for identifying knowledge gaps, as thus as a to guide future research.
We illustrate our approach using two examples, and demonstrate that the framework can reduce bias and uncertainty when projecting beyond the range of the original data.
In order to provide a platform for continuing to develop the framework, we include all code and data for the examples as supporting information with the manuscript.
Understanding range shifts are of broad interest for both theoretical and applied ecologists, and we believe that our manuscript represents an important step toward a more holistic approach for forecasting species ranges.
We hope you will consider it for the \emph{Ideas and Perspectives} section. 
We will be ready to submit the manuscript within one month.

%TC:break closing
\closing{Sincerely,}

\end{letter}
\end{document}